---
title: "Assignment_1"
author: "Tirumal Achina"
date: "2023-09-07"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

#Importing the data.
#Source: chatGpt and github
#Downloaded the dataset and read it from my desktop.
```{r}
titanic <- read.csv("C:/Users/tirum/Desktop/titanic/titanic.csv")
#titanic
#the above line would give the whole dataset.
```

#Descriptive analysis: This shows the summary of whole dataset that we have given. It includes the Distribution, Measures of Central Tendency(Mean,Median,Mode),Measures of Variability(Standard Deviation,Range,Variance,Interquartile Range).

```{r}
head(titanic) #This function shows the frst 6 rows and columns of the dataset
tail(titanic) #This function shows the last 6 rows and columns of dataset.
summary(titanic) #This shows the whole summary that includes mean, median,mode and so on.

```
#Transforming variables: This can be done in variable ways such as log,exponential,normalization and so on...For this dataset I choosed Normalization Transformation.
#Normalization Transformation: This changes the scale of variable where the mean becomes 0 and the standard deviation becomes 1. It can be done by min-max or z-score normalization.I have Normalized Fare column in the titanic table.

```{r}
min_max_normalize <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}
#Here the Age got normalized and we can find that values in the Age_Norm column
titanic$Age_norm <- min_max_normalize(titanic$Age)
head(titanic)

```
```{r}
#Log Transformation to Fare Column.
titanic$Fare_log = log(titanic$Fare)
titanic$Fare_log
```

#Hist plots

```{r}
hist(titanic$Age, xlab = "Age", ylab = "Frequency",main = "Hist_Age",col = "white")

#plot(titanic$Age, xlab = "Age",ylab = "Frequency",main = "Hist_Age)
```


#Marking the Scatter plot for Age and Fare

```{r}

x = titanic$Age #Age on the x-axis
y = titanic$Fare #Fare on the y-axis
plot(x,y, xlab ="Age",ylab = "Fare",main = "Age-Fare Plot",pch = 19,col =  "cyan")

```

```{r}
x = titanic$Age
y = titanic$Pclass
plot(x,y, xlab ="Age",ylab = "Pclass",main = "Age-Pclass Plot",pch = 19,col =  "cyan")
```

